\acknowledge{
%All of this work would not have been possible without the guidance of my advisor Eric Corwin. Special thanks go to Varda Faghir Hagh, the postdoctoral associate in my research lab, who I worked closely with for my latest project. I would like to express my gratitude to Giulio Biroli and Ludovic Berthier for giving me the opportunity of participating to the writing of the review article present in this dissertation, and to Fran\c cois Landes for a thriving peer collaboration. For various insights and support, I would like to thank my lab mates: James Sartor, Valerie Beale, Aileen Carroll-Godfrey, Cameron Dennis, Andrew Hammond and Jacob Hass. I would also like to thank my committee for their advice at various points throughout my graduate career.
Full acknowledgement of all the support I've received through the writing of this would be longer than the dissertation itself. Nevertheless, I'd like to take a moment to express my gratitude here to those who have contributed the most to this work.

First and foremost I would like to thank my advisor Eric ``Irwin'' Corwin, who provided the funding, drive, motivation, research question, expertise, dogs, etc.

Special thanks to Sean Ridout for not only the contributions to the letter which we wrote together, but also for helpful feedback on my other projects. 
I would also like to acknowledge Francesco Zamponi for graciously assisting us in interpreting our results in the same aforementioned letter.

I'd also like to thank the labmates I have had during my time here for myriads of help and support: Cam Dennis, Francesco Arceri, Varda Faghir Hagh, Andy Hammond, Valerie Beale, Aileen Carroll-Godfrey, Jack Dale, Jacob Hass, Alex Trevelyan, Georgios Tsekenis, and Mike van der Naald.

Thanks to CorwinLab alumni Peter Morse for advice on both research and death cults, and Kyle Welch for getting us into PRL.

I'd also like to thank all my friends and family for their support through this time, as well as innumerable fellow grad students, faculty, and staff who have helped me in various ways throughout this time.

This work was supported by the NSF Career Award grant No. DMR-1255370, and the Simons Collaboration on Cracking the Glass Problem (No. 454939 E. Corwin)}
